---
title: "Results_Dec2022"
author: "Caitie"
date: "2022-12-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(countrycode)
```

# Load data

```{r}
yearly_trade<-read.csv(here("output_data/yearly_trade_results.csv"))
trade_data<-read.csv(here("output_data/Trade_data_clean_all.csv"))
trade_data_spp<-read.csv(here("output_data/Trade_data_clean_speciesonly.csv"))
```

# Results
 
 How much was traded between 2000-2015
 
```{r}
min(yearly_trade$global_trade, na.rm = T)/1000000
max(yearly_trade$global_trade, na.rm = T)/1000000
total_trade<-sum(yearly_trade$global_trade, na.rm = T)/1000000 #Megatonnes
total_trade
```

Amount identified to species level


```{r}
min(yearly_trade$species_trade, na.rm = T)/1000000
max(yearly_trade$species_trade, na.rm = T)/1000000
spp_trade<-sum(yearly_trade$species_trade, na.rm = T)/1000000 #Megatonnes
spp_trade
```




```{r}
unknown_spp<-yearly_trade %>% 
  mutate(perc_spp_not_id = 1-(species_trade/global_trade))

min(unknown_spp$perc_spp_not_id)
max(unknown_spp$perc_spp_not_id)

min(unknown_spp$species_trade, na.rm = T)/1000000
max(unknown_spp$species_trade, na.rm = T)/1000000

# Total percent unknown spp across the entire time period
sum(unknown_spp$species_trade)/sum(yearly_trade$global_trade, na.rm = T)
```

Two-way trade
```{r}
min(yearly_trade$total_like, na.rm = T)/1000000
max(yearly_trade$total_like, na.rm = T)/1000000

like_trade<-yearly_trade %>% 
  mutate(prop_like = total_like/global_trade,
         prop_like_spp = total_like/species_trade)

# Of total global trade
min(like_trade$prop_like, na.rm = T)
max(like_trade$prop_like, na.rm = T)

# Of trade identified to species
min(like_trade$prop_like_spp, na.rm = T)
max(like_trade$prop_like_spp, na.rm = T)
```
Redundant two-way trade
```{r}
min(yearly_trade$total_redundant, na.rm = T)/1000 # note thousands
max(yearly_trade$total_redundant, na.rm = T)/1000

redundant_trade<-yearly_trade %>% 
  mutate(prop_redundant = total_redundant/global_trade,
         prop_redundant_spp = total_redundant/species_trade)

min(redundant_trade$prop_redundant, na.rm = T)
max(redundant_trade$prop_redundant, na.rm = T)

mean(yearly_trade$total_redundant, na.rm = T)/1000
mean(redundant_trade$prop_redundant, na.rm = T)
```

```{r}
# percent increase

(562617.3- 392372.5)/392372.5
```

```{r}
# trend

mod1<-lm(yearly_trade$total_redundant ~ yearly_trade$year)

mod1

```

```{r}
sum_global<-sum(yearly_trade$global_trade,na.rm = T)/1000000
sum_species<-sum(yearly_trade$species_trade, na.rm = T)/1000000
sum_like<-sum(yearly_trade$total_like,na.rm = T)/1000000
sum_redundant<-sum(yearly_trade$total_redundant,na.rm = T)/1000000

sum_like/sum_global
sum_redundant/sum_global
sum_redundant/sum_species
```

# Country results

```{r}
trade_data<-read.csv(here("output_data/Trade_data_clean_all.csv"))
```

All countries that traded
```{r}
exp_iso3<-unique(trade_data$ExportISO3, na.rm = T)
imp_iso3<-unique(trade_data$ImportISO3, na.rm = T)

all_iso3<-c(exp_iso3, imp_iso3)

length(unique(all_iso3))
```

```{r}
redun<-read.csv(here("output_data/Total_redundant_trade_by_exportISO3.csv"))
```


```{r}
redun<-redun %>% 
  arrange(desc(Redun_tonnes)) %>% 
  mutate(cumsum = cumsum(Redun_tonnes)/1000000,
         prop_cumsum = cumsum/sum_redundant)
```

```{r}
cor.test(redun$Export_tonnes, redun$Redun_tonnes, method = "pearson")
```

```{r}
redun_50<-redun %>% 
  filter(prop>=0.5)

redun_100<-redun %>% 
  filter(prop==1) %>% 
  mutate(Country = countrycode(ExportISO3, "iso3c", "country.name"))

length(which(redun$prop>=0.5))
length(which(redun$prop==1))

(sum(redun_100$Redun_tonnes))
(sum(redun_100$Redun_tonnes)/1000000)/sum_redundant
```

# Trade partners

```{r}
partners<-read.csv(here("output_data/Country_redundant_trade_combinations.csv"))
```

```{r}
partners<-partners %>% 
  arrange(desc(redun_like_tonnes)) %>% 
  mutate(cumsum = cumsum(redun_like_tonnes)/1000000,
         prop_cumsum = cumsum/sum_redundant,
         prop = (redun_like_tonnes/1000000)/sum_redundant)

length(which(partners$prop_cumsum<=0.9))
```

```{r}
num_partners<-read.csv(here("output_data/Number_Trade_Partners.csv"))

length(which(num_partners$trade_partners == 1))
```

```{r}
cont<-read.csv(paste0(here("raw_data/Countries_by_continent.csv"))) %>% 
  dplyr::select(ISO3, Region1, Continent)

cont_partners<-partners %>% 
  left_join(., cont, by = c("x" = "ISO3")) %>% 
  left_join(., cont, by = c("y" = "ISO3"))

cont_partners<-cont_partners %>% 
  mutate(Continent.y = ifelse(y == "TWN", "Asia", Continent.y),
         Region1.y = ifelse(y == "TWN", "Eastern Asia", Region1.y))

length(which(cont_partners$Continent.x == cont_partners$Continent.y))
length(which(cont_partners$Continent.x != cont_partners$Continent.y))

length(which(cont_partners$Region1.x == cont_partners$Region1.y))
length(which(cont_partners$Region1.x != cont_partners$Region1.y))
```

```{r}
diff_cont<-cont_partners %>% 
  filter(Continent.x != Continent.y)
```

# China

Look into China

```{r}
test<-partners %>%  filter(x == "CHN" | y == "CHN")
```

```{r}
test2<-trade_data %>% 
  filter(ExportISO3 == "CHN" | ImportISO3 == "CHN")

test3<-trade_data %>% 
  filter(ExportISO3 == "CHN")
```

```{r}
test4<-trade_data_spp %>% 
  filter(ExportISO3 == "CHN")

```

# Denmark and Germany

```{r}
deu<-partners %>% 
  filter(x == "DEU" | y == "DEU")

deu<-deu %>% 
  mutate(prop = (redun_like_tonnes/2)/(sum(deu$redun_like_tonnes)/2))

dnk<-partners %>% 
  filter(x == "DNK" | y == "DNK") 

dnk<-dnk %>% 
  mutate(prop = (redun_like_tonnes/2)/(sum(dnk$redun_like_tonnes)/2))
```

```{r}
deu_dnk<-read.csv(here("output_data/Country_Country/DEU_DNK_like_species_trade.csv"))
```

```{r}
deu_dnk_spp<-deu_dnk %>% 
  group_by(CommonName, TaxonName) %>% 
  summarise(total_tonnes = sum(Tonnage, na.rm = T))
```
# Species results

All species traded (identified to species level)

```{r}
length(unique(trade_data_spp$TaxonName))
```
Total countries with species info

```{r}
exportiso3<-unique(trade_data_spp$ExportISO3)
importiso3<-unique(trade_data_spp$ImportISO3)

trade_spp_iso3<-c(exportiso3, importiso3)

length(unique(trade_spp_iso3))
```


```{r}
spp_dat<-read.csv(here("output_data/Country_redundant_species_results.csv"))

spp_dat2<-spp_dat %>% 
  group_by(TaxonName, CommonName) %>% 
  summarise(total_like = sum(like_tonnes, na.rm = T),
            total_redun = sum(redun_tonnes, na.rm = T))

spp_dat2<-spp_dat2 %>% 
  arrange(desc(total_redun))

length(which(spp_dat2$total_redun>0))

```

```{r}
spp_dat2 <-spp_dat2 %>% 
  arrange(desc(total_redun)) %>% 
  ungroup() %>% 
  mutate(cumsum= cumsum(total_redun)/1000000,
         prop_cumsum = cumsum/sum_redundant,
         prop = (total_redun/1000000)/sum_redundant)

length(which(spp_dat2$prop_cumsum<=0.9))
```

```{r}
num_redun_sp<-read.csv(here("output_data/Number_of_species_traded_by_country.csv"))
```

```{r}
options(scipen = 999)
sub_spp<-spp_dat %>% 
  filter(CommonName %in% c("Atlantic herring", "Skipjack tuna", "Atlantic cod", "Atlantic mackerel")) %>%
  group_by(c1, c2, CommonName) %>% 
  summarise(total_redun = sum(redun_tonnes, na.rm = T))



ah<-sub_spp %>% 
  filter(CommonName == "Atlantic herring") %>% 
  mutate(prop = total_redun/2380877.34172)

ac<-sub_spp %>% 
  filter(CommonName == "Atlantic cod") %>% 
  mutate(prop = total_redun/724473.57872)

sj<-sub_spp %>% 
  filter(CommonName == "Skipjack tuna") %>% 
  mutate(prop = total_redun/619533.39460)

am<-sub_spp %>% 
  filter(CommonName == "Atlantic mackerel")%>% 
  mutate(prop = total_redun/591269.95871)
```


# Proportion of trade identified to species for each country

```{r}
all_trade<-trade_data %>% 
  group_by(ExportISO3, ExportCountry) %>% 
  summarise(all_trade = sum(Tonnage, na.rm = T))

all_trade_spp<-trade_data_spp %>% 
  group_by(ExportISO3) %>% 
  summarise(spp_trade = sum(Tonnage, na.rm = T))

join_trade<-full_join(all_trade, all_trade_spp) %>% 
  mutate(spp_trade = ifelse(is.na(spp_trade) == T, 0, spp_trade),
    prop_spp_trade = spp_trade/all_trade)
```

```{r}
write.csv(join_trade, here("output_data/Prop_trade_identified_to_species_by_country.csv"))
```


```{r}
test<-left_join(redun, join_trade)

cor.test(test$Redun_tonnes, test$prop_spp_trade)
```

